Estimating Access to Rabies Post-Exposure Prophylaxis in Madagascar

### 2019-04-02 accessPEP.Rmd
fixed vial analyses and running distance models for burden analyses, etc.

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
## Clear memory
rm(list = ls())
## Libraries
library(data.table)
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
library(stringr)
library(ISOweek)
library(raster)
library(sp)
library(MASS)
library(reshape2)
library(geosphere)
library(foreach)
library(bbmle)
library(boot)
library(geosphere)
library(iterators)
library(patchwork)

## Files
source("R/utils.R")
source("R/models.R")
source("R/data_functions.R")
source("R/get.burden.R")
select <- dplyr::select
knitr::opts_chunk$set(fig.pos = 'h', echo = FALSE, message = FALSE, warning = FALSE, 
                      include = TRUE)
```

@@

figM1

@@

For most districts, the majority of bites were reported to the closest clinic as estimated by our travel time metric (Fig 1A). Therefore, we defined the catchment area for each clinic as all districts for which the clinic was the closest ARMC. We excluded any catchment areas for which the clinic did not submit any forms (catchments = `r length(no_data)`, districts = `r nrow(exps_dist) - nrow(mada_district)`, in grey in Fig 1A).

@@

### Model of reported bites as a function of travel time
We modeled incidence of bites as a function of travel times in hours as follows:

\begin{equation}
  \mu = exp(\beta_{t}T + \beta_0)P 
\begin{equation}

@@

\begin{equation}
  \mu = \sum \limits_{j=1}^jexp(\beta_{t}T_j + \beta_0)P_j 
\begin{equation}

@@

\begin{equation}
  \overline\rho = \frac{\mu p_{rabid}}{R}
\begin{equation}

@@

\begin{equation}
           p_{rabid}=
            \begin{cases}
            x, & \text{if}\ \frac{R\rho_{max}}{B} > x \\
            \frac{R\rho_{max}}{B}, & \text{otherwise}
            \end{cases}
\begin{equation}
<br>

@@

\begin{equation}
           \rho_d = \sum \limits_{j=1}\frac{rho_j*P_j}{P_d}
\begin{equation}

@@

\begin{equation}
deaths_i = (R_i - p_{rabid_i}B_i) \times p_{death}
\begin{equation}

@@

### Sensitivity analyses for burden estimates 
We estimate burden deterministically across a range of parameter values to test the effects of our model assumptions on estimates of rabies burden. Specifically, we fix rabies incidence at the minimum and maximum of our estimated range, look at the range of values of p_rabid between 0.2 - 0.6 (as per Moramanga), and the range of values of rho_max (0.85 - 0.99) to get at maximum and minimum estimates of burden. We also examine the impact of systematic variation in rabies incidence with human population (a potential proxy for changes in the dog population, see section S5) by looking at how estimates of burden change if rabies incidence were to scale positively or negatively with human population size.
<br>

@@

### Models of bites as a function of travel times
```{r fig M2 models of ttimes, cache = TRUE}
## Exclude for comm data!
comm_covars <- filter(comm_data, exclude == 0)

### Running estimation
mada_mods_covars <- run.mods(bites = exps_dist$bites, names_bites = exps_dist$district,
                             pop_predict = c("flatPop"),
                             pop_vars = list(pop_dist = exps_dist$pop, pop_commune = comm_covars$pop),
                             covars = list(ttimes_district = exps_dist$ttimes_weighted_dist/60,
                                           ttimes_commune = comm_covars$ttimes_weighted/60), 
                             sum_it = c(FALSE, TRUE, FALSE, TRUE), run_type = "optim", label = "Mada")
mora_mods_covars <- run.mods(bites = exps_mora$bites, names_bites = exps_mora$commune,
                             pop_predict = c("flatPop"),
                             pop_vars = list(pop_dist = exps_mora$pop),
                             covars = list(ttimes_mora = exps_mora$ttimes_weighted/60), 
                             sum_it = c(FALSE, FALSE), run_type = "optim", label = "Mora")
mada_mods <- c(mada_mods_covars, mora_mods_covars)

@@

preds_fixed <- run.predicts.fixed(covar_df = bite_mods, pop_fixed = 1e5, pop_seq = pop_plot)
preds_data <- run.predicts.data(covar_df = bite_mods, run_type = "predict.data")
preds_raw <- run.predicts.data(covar_df = bite_mods, run_type = "predict.raw")

a <- ggplot(preds_fixed, aes(x = covar, y = predicted, color = scale)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = scale), color = NA, alpha = 0.5) +
  scale_color_manual(values = c("navy", "purple", "red"), name = "Model scale") +
  scale_fill_manual(values = c("blue", "lavender", "pink"), guide = "none") +
  xlab("Travel time (hrs)") +
  ylab("Reported bites per \n 100,000 persons") +
  labs(tag = "A")

b <- ggplot(preds_data, aes(x = observed, y = predicted, color = scale)) + 
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey") +
  expand_limits(y = c(0, 4000)) +
  scale_color_manual(values = c("navy", "purple", "red"), name = "Model scale") +
  xlab("Mean annual bites observed") +
  ylab("Mean annual bites predicted") +
  labs(tag = "B")

figM2 <- a / b
figM2

@@

kable(bite_mods)

@@

We estimated similar parameter values from our commune-level data from the Moramanga ARMC and the district level data from 19 clinics across the country (Table 1, Figure 2A), with reported bite incidence decreasing with travel times to the ARMC. The model of bites as a function of travel times at the commune level summed to the district level resulted in a fit closer to the model fit to the Moramanga data at the commune level than the district model. All of the models produced reasonable fits to the data (Figure 5B), however, there was some variation in bite incidence that was not captured by the models, generally underestimating bite incidence in some districts.

### Estimation of burden and reporting 
```{r fig M3 reporting}
B_ttimes_mada <- bite_mods$beta[bite_mods$model == "Mada flatPop ttimes_commune"]
B_0_mada <- bite_mods$intercept[bite_mods$model == "Mada flatPop ttimes_commune"]

B_ttimes_dist <- bite_mods$beta[bite_mods$model == "Mada flatPop ttimes_district"]
B_0_dist <- bite_mods$intercept[bite_mods$model == "Mada flatPop ttimes_district"]

@@

rep_df <- 
  foreach(i = 1:length(p_rabid_vals), .combine = rbind) %:% 
  foreach(k = 1:length(hdr_vals), .combine = rbind) %:%
  foreach(j = 1:length(beta_vals), .combine = rbind) %do% {
    rep <- as.data.frame(list(p_rabid = p_rabid_vals[i], ttimes = ttimes, hdr = hdr_vals[k], 
                                         reporting = get.burden.fixed(pop = 1e5, ttimes = ttimes,
                                                          B_ttimes = beta_vals[[j]], B_0 = int_vals[[j]],
                                            hdr = hdr_vals[k], p_rabid = p_rabid_vals[i],
                                            type = "reporting", scale = FALSE), scale = names(beta_vals)[j]))
  }

rep_min_max <- 
  foreach(j = 1:length(beta_vals), .combine = rbind) %do% {
    rep_max <- get.burden.fixed(pop = 1e5, ttimes = ttimes, B_ttimes = beta_vals[[j]], B_0 = int_vals[[j]],
                                            hdr = 25, p_rabid = 0.6,
                                            type = "reporting", scale = FALSE)

    rep_min <- get.burden.fixed(pop = 1e5, ttimes = ttimes, B_ttimes = beta_vals[[j]], B_0 = int_vals[[j]],
                                            hdr = 5, p_rabid = 0.2,
                                            type = "reporting", scale = FALSE)
    rep <- as.data.frame(list(ttimes = ttimes, rep_max = rep_max, rep_min = rep_min, 
                              scale = names(beta_vals)[j]))
  }

# ggplot(rep_df, aes(x = ttimes, y = reporting, color = interaction(p_rabid, hdr), linetype = as.factor(hdr))) +
#   geom_line() +
#   facet_wrap(~ scale) +
#   scale_color_manual(values = c("red", "red", "blue", "blue"))

c <- ggplot(rep_min_max, aes(x = ttimes, y = rep_min, fill = scale)) +
  geom_line(aes(y = rep_min, color = scale), linetype = 2) +
  geom_line(aes(y = rep_max, color = scale), linetype = 1) +
  geom_ribbon(aes(ymax = rep_max, ymin = rep_min), alpha = 0.5) +
  scale_fill_manual(values = c("purple", "navy")) +
  scale_color_manual(values = c("purple", "navy")) +
  ylab("Reporting") +
  xlab("Travel times (in hours)") +
  labs(tag = "A")

@@

rep_df <- 
  foreach(i = 1:length(p_rabid_vals), .combine = rbind) %:% 
    foreach(j = 1:length(hdr_vals), .combine = rbind) %do% {
    reporting_comm <- get.burden.fixed(pop = comm_data$pop, ttimes = comm_data$ttimes_weighted/60, 
                                            B_ttimes = B_ttimes_mada, B_0 = B_0_mada,
                                            hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], 
                                      type = "reporting", scale = FALSE)
    comm <- as.data.frame(list(names = comm_data$mdg_dis_co, reporting = reporting_comm, hdr = hdr_vals[j], 
                               pop = comm_data$pop, ttimes = comm_data$ttimes_weighted,
                               p_rabid = p_rabid_vals[i], 
                               scale = "Commune", params = "Commune model"))

    reporting_dist <- get.burden.fixed(pop = dist_data$pop, ttimes = dist_data$ttimes_weighted_dist/60, 
                                            B_ttimes = B_ttimes_dist, B_0 = B_0_dist,
                                            hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], 
                                      type = "reporting", scale = FALSE)

     dist <- as.data.frame(list(names = dist_data$district, reporting = reporting_dist, 
                                pop = dist_data$pop, ttimes = dist_data$ttimes_weighted_dist,
                                hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], scale = "District",
                                params = "District model"))
     check <- bind_rows(comm, dist)
  }

d <- ggplot(reporting, aes(x = reorder(names, ttimes), y = reporting, color = scale)) +
  geom_boxplot(outlier.shape = NA) +
  scale_color_manual(values = c(alpha("blue", 0.5), alpha("purple", 0.5)), name = "Parameters") +
  xlab("Districts ordered by travel times \n (low to high)") +
  ylab("Reporting of rabid exposures") +
  theme(axis.text.x = element_blank()) +
  labs(tag = "B")

figM3 <- c / d 
figM3 
ggsave("figs/M3.jpeg", figM3, device = "jpeg", width = 7, height = 7)

@@

```{r run burden and scenario, cache = TRUE}
## District travel times
dist_ttimes <- read.csv("output/district_temp_scenario_20190212_181056.csv", row.names = 1)
comm_ttimes <- read.csv("output/commune_temp_scenario_20190212_182810.csv", row.names = 1)

dist_ttimes <- cbind(dist_catchments$ttimes_weighted, dist_ttimes)
dist_mat <- matrix(NA, nrow(dist_ttimes), ncol(dist_ttimes))
dist_mat[, 1] <- dist_ttimes[, 1]
for (i in 2:ncol(dist_ttimes)){
  dist_mat[, i] <- ifelse(dist_ttimes[, i] < dist_ttimes[, i - 1], dist_ttimes[, i], NA)
}
dist_mat <- as.data.table(dist_mat)
dist_mat$mdg_dis_co <- as.character(mada_district$mdg_dis_co)
dist_mat <- melt(dist_mat, id.vars = "mdg_dis_co", value.name = "ttimes", variable.name = "scenario")
dist_mat$scenario <- as.numeric(gsub("V", "", dist_mat$scenario)) - 1

@@

figM5 <- e / f
ggsave("figs/M5.jpeg", figM5, device = "jpeg")

@@

figS2.1 <- A + B
ggsave("figs/S2.1.jpeg", figS2.1, device = "jpeg")
```

>>>>

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
## Clear memory
rm(list = ls())
## Libraries
library(data.table)
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
library(stringr)
library(ISOweek)
library(raster)
library(sp)
library(MASS)
library(reshape2)
library(geosphere)
library(foreach)
library(bbmle)
library(boot)
library(geosphere)
library(iterators)
library(patchwork)
library(doParallel)

## Files
source("R/utils.R")
source("R/models.R")
source("R/data_functions.R")
source("R/get.burden.R")
select <- dplyr::select
knitr::opts_chunk$set(fig.pos = 'h', echo = FALSE, message = FALSE, warning = FALSE, 
                      include = TRUE, cache.lazy = FALSE)
```

@@

## figM1

@@

For most districts, the majority of bites were reported to the closest clinic as estimated by our travel time metric (Fig 1A). Therefore, we defined the catchment area for each clinic as all districts for which the clinic was the closest ARMC. We excluded any catchment areas for which the clinic did not submit any forms (catchments = `r length(no_data)`, districts = `r nrow(mada_district) - nrow(exps_dist)`, in grey in Fig 1A).

@@

### Model of reported bites as a function of travel time
We modeled incidence of bites as a function of travel times in hours as follows:
$$
  \mu = exp(\beta_{t}T + \beta_0)P 
$$

@@

$$
  \mu = \sum \limits_{j=1}^jexp(\beta_{t}T_j + \beta_0)P_j 
$$

@@

$$
  \overline\rho = \frac{\mu p_{rabid}}{R}
$$

@@

$$
           p_{rabid}=
            \begin{cases}
            x, & \text{if}\ \frac{R\rho_{max}}{B} > x \\
            \frac{R\rho_{max}}{B}, & \text{otherwise}
            \end{cases}
$$
<br>

@@

$$
           \rho_d = \sum \limits_{j=1}\frac{rho_j*P_j}{P_d}
$$

@@

$$
deaths_i = (R_i - p_{rabid_i}B_i) \times p_{death}
$$

@@

### Sensitivity analyses for burden estimates 
We estimate burden deterministically across a range of parameter values to test the effects of our model assumptions on estimates of rabies burden. Specifically, we fix rabies incidence at the minimum and maximum of our estimated range, look at the range of values of $p_{rabid}$ between 0.2 - 0.6 (as per Moramanga), and the range of values of $rho_{max}$ (0.85 - 0.99) to get at maximum and minimum estimates of burden. We also examine the impact of systematic variation in rabies incidence with human population (a potential proxy for changes in the dog population, see section S5) by looking at how estimates of burden change if rabies incidence were to scale positively or negatively with human population size.
<br>

@@

### Models of bites as a function of access
```{r fig M2 models of ttimes, cache = TRUE}
## Exclude for comm data!
comm_covars <- filter(comm_data, exclude == 0)

### Running estimation
mada_mods_covars <- run.mods(bites = exps_dist$bites, names_bites = exps_dist$district,
                             pop_predict = c("flatPop"),
                             pop_vars = list(pop_dist = exps_dist$pop, 
                                             pop_commune = comm_covars$pop),
                             covars = list(distance_district = exps_dist$distance, 
                                           distance_commune = comm_covars$distance,
                                           ttimes_district = exps_dist$ttimes_weighted_dist/60,
                                           ttimes_commune = comm_covars$ttimes_weighted/60), 
                             sum_it = c(FALSE, TRUE, FALSE, TRUE), 
                             run_type = "optim", label = "Mada")

mora_mods_covars <- run.mods(bites = exps_mora$bites, names_bites = exps_mora$commune,
                             pop_predict = c("flatPop"),
                             pop_vars = list(pop_dist = exps_mora$pop),
                             covars = list(distance_mora = exps_mora$distance,
                                           ttimes_mora = exps_mora$ttimes_weighted/60), 
                             sum_it = c(FALSE, FALSE), run_type = "optim", label = "Mora")
mada_mods <- c(mada_mods_covars, mora_mods_covars)

@@

preds_fixed <- run.predicts.fixed(covar_df = bite_mods, pop_fixed = 1e5, pop_seq = pop_plot)
preds_data <- run.predicts.data(covar_df = bite_mods, run_type = "predict.data")
preds_raw <- run.predicts.data(covar_df = bite_mods, run_type = "predict.raw")
covar_labels <- c("distance" = "Distance (km)", "ttimes" = "Travel times (hrs)")
a <- ggplot(preds_fixed, aes(x = covar, y = predicted, color = scale)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = scale), color = NA, alpha = 0.45) +
  scale_color_manual(values = c("navy", "purple", "red"), name = "Model scale") +
  scale_fill_manual(values = c("blue", "purple", "pink"), guide = "none") +
  xlab("Access metric") +
  ylab("Reported bites per \n 100,000 persons") +
  labs(tag = "A") + 
  facet_wrap(~covar_name, scales = "free_x", 
             labeller = labeller(covar_name = as_labeller(covar_labels)))

b <- ggplot(preds_data, aes(x = observed, y = predicted, color = scale)) + 
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey") +
  expand_limits(y = c(0, 4000)) +
  scale_color_manual(values = c("navy", "purple", "red"), name = "Model scale") +
  xlab("Mean annual bites observed") +
  ylab("Mean annual bites predicted") +
  facet_wrap(~covar_name, scales = "free_x", 
             labeller = labeller(covar_name = as_labeller(covar_labels))) +
  labs(tag = "B")

figM2 <- a / b
## figM2

@@

kable(bite_mods)
bite_mods %>%
  mutate(model_scale = 
           case_when(grepl("district", model, fixed = TRUE) ~ "District",
                     grepl("commune", model, fixed = TRUE) ~ "Commune", 
                     grepl("mora", model, fixed = TRUE) ~ "Moramanga"),
         model_covar = 
           case_when(grepl("distance", model, fixed = TRUE) ~ "Distance (km)",
                     grepl("ttimes", model, fixed = TRUE) ~ "Travel time (hrs)")) -> bite_mods

beta_vals <- paste0(round(bite_mods$beta, 3), 
               paste0(" (", round(bite_mods$beta_lower, 3), " - ", round(bite_mods$beta_upper, 3), ")"))
intercept_vals <- paste0(round(bite_mods$intercept, 3), 
               paste0(" (", round(bite_mods$intercept_lower, 3), " - ", round(bite_mods$intercept_upper, 3), ")"))
likelihood <- round(bite_mods$likelihood, 2)
kable(cbind(bite_mods$model_covar, bite_mods$model_scale, beta_vals, intercept_vals, likelihood), 
      col.names = c("Covariate", "Model", "Access effect", "Intercept", "Likelihood"))

@@

We estimated similar parameter values from our commune-level data from the Moramanga ARMC and the district level data from 19 clinics across the country (Table 1, Figure 2A), with reported bite incidence decreasing with increased distance or travel times to the ARMC. The models of bites as a function of access at the commune level summed to the district level resulted in a fit closer to the model fit to the Moramanga data at the commune level than the district model. All of the models produced reasonable fits to the data (Figure 2B), however there were multiple districts for the models underestimated observed bites from the national data (points below the dashed grey 1:1 line in Figure 2B), although the Commune model seemed to better capture observed bites than the District model in these cases.

### Estimation of burden and reporting 
```{r fig M3 reporting}
## Params from distance models
B_distance_mada <- bite_mods$beta[bite_mods$model == "Mada flatPop distance_commune"]
B_0_distance_mada <- bite_mods$intercept[bite_mods$model == "Mada flatPop distance_commune"]
B_distance_dist <- bite_mods$beta[bite_mods$model == "Mada flatPop distance_district"]
B_0_distance_dist <- bite_mods$intercept[bite_mods$model == "Mada flatPop distance_district"]

## Params from ttime models
B_ttimes_mada <- bite_mods$beta[bite_mods$model == "Mada flatPop ttimes_commune"]
B_0_mada <- bite_mods$intercept[bite_mods$model == "Mada flatPop ttimes_commune"]
B_ttimes_dist <- bite_mods$beta[bite_mods$model == "Mada flatPop ttimes_district"]
B_0_dist <- bite_mods$intercept[bite_mods$model == "Mada flatPop ttimes_district"]

@@

rep_df_ttimes <- 
  foreach(i = 1:length(p_rabid_vals), .combine = rbind) %:% 
  foreach(k = 1:length(hdr_vals), .combine = rbind) %:%
  foreach(j = 1:length(beta_vals), .combine = rbind) %do% {
    rep <- as.data.frame(list(p_rabid = p_rabid_vals[i], covar = ttimes, hdr = hdr_vals[k],
                              reporting = get.burden.fixed(pop = 1e5, covar = ttimes,
                                                          B_covar = beta_vals[[j]], 
                                                          B_0 = int_vals[[j]],
                                                          hdr = hdr_vals[k], 
                                                          p_rabid = p_rabid_vals[i],
                                                          type = "reporting", scale = FALSE), 
                              model = names(beta_vals)[j]))
  }
rep_df_ttimes$covar_name <- "Travel times (hrs)"
rep_min_max_ttimes <- 
  foreach(j = 1:length(beta_vals), .combine = rbind) %do% {
    rep_max <- get.burden.fixed(pop = 1e5, covar = ttimes, 
                                B_covar = beta_vals[[j]], B_0 = int_vals[[j]],
                                            hdr = 25, p_rabid = 0.6,
                                            type = "reporting", scale = FALSE)

    rep_min <- get.burden.fixed(pop = 1e5, covar = ttimes, B_covar= beta_vals[[j]], 
                                B_0 = int_vals[[j]], hdr = 5, p_rabid = 0.2, 
                                type = "reporting", scale = FALSE)
    rep <- as.data.frame(list(covar = ttimes, rep_max = rep_max, rep_min = rep_min, 
                              model = names(beta_vals)[j]))
  }
rep_min_max_ttimes$covar_name <- "Travel times (hrs)"

beta_vals <- list(District = B_distance_dist, Commune = B_distance_mada)
int_vals <- list(District = B_0_distance_dist, Commune = B_0_distance_mada)
distance <- seq(0, 200, by = 5)

rep_df_distance <- 
  foreach(i = 1:length(p_rabid_vals), .combine = rbind) %:% 
  foreach(k = 1:length(hdr_vals), .combine = rbind) %:%
  foreach(j = 1:length(beta_vals), .combine = rbind) %do% {
    rep <- as.data.frame(list(p_rabid = p_rabid_vals[i], covar = distance, hdr = hdr_vals[k],
                              reporting = get.burden.fixed(pop = 1e5, covar = distance,
                                                          B_covar = beta_vals[[j]], 
                                                          B_0 = int_vals[[j]],
                                                          hdr = hdr_vals[k], 
                                                          p_rabid = p_rabid_vals[i],
                                                          type = "reporting", scale = FALSE), 
                              model = names(beta_vals)[j]))
  }
rep_df_distance$covar_name <- "Distance  (hrs)"

rep_min_max_distance <- 
  foreach(j = 1:length(beta_vals), .combine = rbind) %do% {
    rep_max <- get.burden.fixed(pop = 1e5, covar = distance, 
                                B_covar = beta_vals[[j]], B_0 = int_vals[[j]],
                                            hdr = 25, p_rabid = 0.6,
                                            type = "reporting", scale = FALSE)
    rep_min <- get.burden.fixed(pop = 1e5, covar = distance, B_covar = beta_vals[[j]], 
                                B_0 = int_vals[[j]], hdr = 5, p_rabid = 0.2, 
                                type = "reporting", scale = FALSE)
    rep <- as.data.frame(list(covar = distance, rep_max = rep_max, rep_min = rep_min, 
                              model = names(beta_vals)[j]))
  }
rep_min_max_distance$covar_name <- "Distance (km)"

rep_min_max <- bind_rows(rep_min_max_ttimes, rep_min_max_distance)
c <- ggplot(rep_min_max, aes(x = covar, y = rep_min, fill = model)) +
  geom_line(aes(y = rep_min, color = model), linetype = 2) +
  geom_line(aes(y = rep_max, color = model), linetype = 1) +
  geom_ribbon(aes(ymax = rep_max, ymin = rep_min), alpha = 0.5) +
  scale_fill_manual(values = c("purple", "navy")) +
  scale_color_manual(values = c("purple", "navy")) +
  ylab("Reporting") +
  facet_wrap(~covar_name, scales = "free_x") +
  xlab("Access metric") +
  labs(tag = "A")

@@

rep_df <- 
  foreach(i = 1:length(p_rabid_vals), .combine = rbind) %:% 
    foreach(j = 1:length(hdr_vals), .combine = rbind) %do% {
    comm_ttimes <- get.burden.fixed(pop = comm_data$pop, 
                                        covar = comm_data$ttimes_weighted/60, 
                                        B_covar = B_ttimes_mada, B_0 = B_0_mada,
                                        hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], 
                                        type = "reporting", scale = FALSE)
    comm_ttimes <- as.data.frame(list(names = comm_data$mdg_dis_co, reporting = comm_ttimes, 
                               hdr = hdr_vals[j], 
                               pop = comm_data$pop, covar = comm_data$ttimes_weighted/60,
                               p_rabid = p_rabid_vals[i], 
                               scale = "Commune", params = "Commune model", 
                               covar_name = "Travel times (hrs)"))
    dist_ttimes <- get.burden.fixed(pop = dist_data$pop, 
                                       covar = dist_data$ttimes_weighted_dist/60, 
                                       B_covar = B_ttimes_dist, B_0 = B_0_dist,
                                       hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], 
                                       type = "reporting", scale = FALSE)
    dist_ttimes <- as.data.frame(list(names = dist_data$district, reporting = dist_ttimes, 
                                pop = dist_data$pop, covar = dist_data$ttimes_weighted_dist,
                                hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], scale = "District",
                                params = "District model", covar_name = "Travel times (hrs)"))

    comm_distance <- get.burden.fixed(pop = comm_data$pop, 
                                        covar = comm_data$distance, 
                                        B_covar = B_distance_mada, B_0 = B_0_distance_mada,
                                        hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], 
                                        type = "reporting", scale = FALSE)
    comm_distance <- as.data.frame(list(names = comm_data$mdg_dis_co, reporting = comm_distance, 
                               hdr = hdr_vals[j], 
                               pop = comm_data$pop, covar = comm_data$distance,
                               p_rabid = p_rabid_vals[i], 
                               scale = "Commune", params = "Commune model", 
                               covar_name = "Distance (km)"))
    dist_distance <- get.burden.fixed(pop = dist_data$pop, 
                                       covar = dist_data$distance, 
                                       B_covar = B_distance_dist, B_0 = B_0_distance_dist,
                                       hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], 
                                       type = "reporting", scale = FALSE)
    dist_distance <- as.data.frame(list(names = dist_data$district, reporting = dist_distance, 
                                pop = dist_data$pop, covar = dist_data$distance,
                                hdr = hdr_vals[j], p_rabid = p_rabid_vals[i], scale = "District",
                                params = "District model", covar_name = "Distance (km)"))
    check <- bind_rows(comm_ttimes, dist_ttimes, comm_distance, dist_distance)
  }

rep_df %>%
  group_by(names, hdr, p_rabid, scale, params, covar_name) %>%
  mutate(pop_total = sum(pop, na.rm = TRUE), prop_pop = pop/pop_total) %>%
  summarize(reporting = sum(reporting*prop_pop), covar = mean(covar, na.rm = TRUE)) -> reporting

d <- ggplot(reporting, aes(x = reorder_within(names, by = covar, within = covar_name), 
                           y = reporting, color = scale)) +
  geom_boxplot(outlier.shape = NA) +
  scale_color_manual(values = c(alpha("blue", 0.5), alpha("purple", 0.5)), name = "Parameters") +
  xlab("Districts ordered by access \n (high to low)") +
  ylab("Reporting of rabid exposures") +
  scale_x_reordered() +
  theme(axis.text.x = element_blank()) +
  facet_wrap(~covar_name, scales = "free_x") +
  labs(tag = "B")

figM3 <- c / d 
## figM3 
ggsave("figs/M3.jpeg", figM3, device = "jpeg", width = 7, height = 7)

@@

```{r run baseline burden, cache = TRUE}
# Calculate the number of cores
no_cores <- 3

distance_dist <- get.burden.stochastic(names = dist_data$district, covar = dist_data$distance,
                           pop = dist_data$pop, scenario = 0,
                           param_covar = B_distance_dist,
                           param_intercept = B_0_distance_dist,
                           p_rab_min = 0.2, p_rab_max = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000, cores = no_cores)

distance_dist$model <- "District"
distance_dist$covar_name <- "Distance (km)"

distance_comm <- get.burden.stochastic(names = comm_data$mdg_dis_co, covar = comm_data$distance,
                           pop = comm_data$pop, scenario = 0,
                           param_covar= B_distance_mada,
                           param_intercept = B_0_distance_mada,
                           p_rab_min = 0.2, p_rab_max = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000, cores = no_cores)
distance_comm$model <- "Commune"
distance_comm$covar_name <- "Distance (km)"

### District ttimes
ttimes_dist <- get.burden.stochastic(names = dist_data$district, 
                                     covar = dist_data$ttimes_weighted_dist/60,
                           pop = dist_data$pop, scenario = 0,
                           param_covar = B_ttimes_dist,
                           param_intercept = B_0_dist,
                           p_rab_min = 0.2, p_rab_max = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000, cores = no_cores)
ttimes_dist$model <- "District"
ttimes_dist$covar_name <- "Travel times (hrs)"

ttimes_comm <- get.burden.stochastic(names = comm_data$mdg_dis_co, 
                                       covar = comm_data$ttimes_weighted/60, 
                           pop = comm_data$pop, scenario = 0,
                           param_covar= B_ttimes_mada,
                           param_intercept = B_0_mada,
                           p_rab_min = 0.2, p_rab_max = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000, cores = no_cores)
ttimes_comm$model <- "Commune"
ttimes_comm$covar_name <- "Travel times (hrs)"

results_base <- bind_rows(ttimes_comm, ttimes_dist, distance_comm, distance_dist)
results_base %>%
  group_by(scenario, names, model, covar_name) %>%
  summarize(deaths_mean = sum(deaths_mean, na.rm = TRUE), 
            deaths_upperCI = sum(deaths_upperCI, na.rm = TRUE),
            deaths_lowerCI = sum(deaths_lowerCI, na.rm = TRUE),
            averted_mean = sum(averted_mean, na.rm = TRUE), 
            averted_upperCI = sum(averted_upperCI, na.rm = TRUE),
            averted_lowerCI = sum(averted_lowerCI, na.rm = TRUE), 
            covar = mean(covar, na.rm = TRUE)) -> results_base_summarized
results_base_summarized %>%
  left_join(dist_data, by = c("names" = "district")) %>%
  drop_na(names) -> results_all_base

ggplot(results_all_base, 
       aes(y = deaths_mean/pop*1e5, x = reorder_within(names, by = covar, within = covar_name), 
                              color = model)) +
  geom_pointrange(aes(ymin = deaths_lowerCI/pop*1e5, ymax = deaths_upperCI/pop*1e5), fatten = 2) +
  xlab("Districts ordered by access \n (high to low)") +
  ylab("Annual deaths per 100k persons") +
  coord_flip() +
  scale_color_manual(values = c(alpha("navy", 0.5), alpha("purple", 0.5)), 
                     name = "Scale") +
  scale_x_reordered() +
  theme(axis.text.y = element_blank()) +
  facet_wrap(~covar_name, scales = "free_y") +
  labs(tag = "A") -> d

p <- ggplot() +
  geom_polygon(data = gg_district_burden, aes(x = long, y = lat, group = group, 
                                              fill = deaths_mean/pop*1e5), 
               color = "grey50") +
  geom_point(data = ctar_points, aes(x = LONGITUDE, y = LATITUDE), color = "blue") + 
  scale_fill_gradient(low = "white", high = "red", name = "Annual deaths per 100k \n persons") +
  labs(tag = "B") +
  facet_grid(covar_name~model)

figM4 <- d / p + plot_layout(ncol = 1, height = c(1, 3))

## figM4

ggsave("figs/M4.jpeg", figM4, device = "jpeg", width = 10, height = 15)

kable(results_all_base %>%
  group_by(model, covar_name) %>%
  summarize(averted = sum(averted_mean), 
            averted_upper = sum(averted_upperCI),
            averted_lower = sum(averted_lowerCI), 
            deaths = sum(deaths_mean), 
            deaths_upper = sum(deaths_upperCI),
            deaths_lower = sum(deaths_lowerCI),
            death_vals = paste0(deaths, 
                                paste0(" (", deaths_upper, " - ", deaths_lower, ")")),
            averted_vals = paste0(averted, 
                                paste0(" (", averted_upper, " - ", averted_lower, ")"))) %>%
            select(`Access metric` = covar_name, Model = model, 
                   `Burden of deaths` = death_vals, `Deaths averted by PEP` = averted_vals) ->
  summary_burden, digits = 2)

Overall, we estimate average annual deaths between `r round(min(summary_burden$deaths))` - `r round(max(summary_burden$deaths))` and deaths_averted between `r round(min(summary_burden$averted))` - `r round(max(summary_burden$averted))`. In general, incidence of rabies deaths increases with travel times, but this effect is more pronounced for the district model compared to the commune model. Our sensitivity analyses show, that while there is considerable uncertainty in the number of deaths estimated, the pattern of deaths increasing with travel times is robust for both models (see figure S5.2). Comparing the two models is, the district model generally predicts lower reported bite incidence overall and less steep declines in reporting with higher travel times.

```{r run expanded scenario, cache = TRUE}
## District travel times
dist_ttimes <- read.csv("output/district_temp_scenario_20190212_181056.csv", row.names = 1)
comm_ttimes <- read.csv("output/commune_temp_scenario_20190212_182810.csv", row.names = 1)

dist_ttimes <- cbind(dist_catchments$ttimes_weighted, dist_ttimes)
dist_mat <- matrix(NA, nrow(dist_ttimes), ncol(dist_ttimes))
dist_mat[, 1] <- dist_ttimes[, 1]

for (i in 2:ncol(dist_ttimes)){
  dist_mat[, i] <- ifelse(dist_ttimes[, i] < dist_ttimes[, i - 1], dist_ttimes[, i], NA)
}

dist_mat <- as.data.table(dist_mat)
dist_mat$mdg_dis_co <- as.character(mada_district$mdg_dis_co)
dist_mat <- melt(dist_mat, id.vars = "mdg_dis_co", value.name = "ttimes", variable.name = "scenario")
dist_mat$scenario <- as.numeric(gsub("V", "", dist_mat$scenario)) - 1

@@

figM5 <- e / f
## figM5
ggsave("figs/M5.jpeg", figM5, device = "jpeg")

@@

figS2.1 <- A / B
## figS2.1
ggsave("figs/S2.1.jpeg", figS2.1, device = "jpeg", height = 7, width = 7)
